\vspace{-1.5cm} "STA610 Case Study 1 Team 4 Report

### 2021-03-22 final_report_2.Rmd

---
title: \vspace{-1.5cm} "STA610 Case Study 1 Team 4 Report"
author:
- Cole Juracek (Programmer)
- Lauren Palazzo (Writer)
- Lingyu Zhou (Checker + Coordinator)
- Fan Zhu (Presenter)
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document:
    df_print: paged
fontsize: 11pt
geometry: margin=2cm
---

@@

For recoding date into a continuous variable, `date_num`, we filter out observations dated prior to the year 2010, as that is the year that the StreetRx web site was founded. For the seasonality variable, `date_quarter`, we extract the "quarter" of the year (numbered 1 through 4) as its own variable. Below we see the distributions of these variables:

@@

```{r}
#hist(df$date_num)
knitr::kable(table(df$date_quarter), col.names = c("Quarter","Freq"))
```

@@

We renamed this variable `dose_per_mg` to make its meaning clearer to the reader. This is a continuous variable that nevertheless tends to take on only a few values, as depicted in the histogram. We considered a recoding of this variable into levels, but decided against it as the dose per milligram technically may take on any value and the generalizability of our model may be improved if we are able to consider dose levels that are in a realistic range yet not present in our current data.

```{r}
hist(df$mgstr)
# rename variable for ease of interpretability
df <- df %>% rename(dose_per_mg = mgstr)
```

@@

To consider whether we may want to include random slopes by state for our predictor variables, we examined plots of the predictor versus response among each state individually to determine whether it seemed that slopes fit to the group-level data would differ. (We present a sample of visualizations here rather than visualizations for every state, due to space constraints.)

For out continuous variables (continuous date and dosage strength), in by-state plots we often saw strikingly similar patterns in predictor vs. response, suggesting that random slopes may not be necessary for these variables.

For categorical variables, we generally observed was there was indeed variation among the states, but it was not clear whether this variation had any well-defined structure. It was not clear whether our models would benefit from random slopes for these variables, so we decided to further investigate this in our modeling process.

@@

## Appendix
```{r, ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}
# Example visualization shown here are for all cities in the state of Florida.

ggplot(df[df$state=="Florida",], aes(x=log_ppm, y=reorder(city, city, length))) +
geom_boxplot() +
labs(title="log PPM by city (FL)", x="Log PPM",y="city")

# More EDA: slope and intercept distributions for predictors by state

# To get a better sense of how state-level groups and slopes are distributed, we created the following intercepts for our coninuous variables:

@@

hist(group_int_list)
hist(group_slope_list)

>>>>

---
title: \vspace{-1.5cm} "STA610 Case Study 1 Team 4 Report"
author:
- Cole Juracek (Programmer)
- Lauren Palazzo (Writer)
- Lingyu Zhou (Checker + Coordinator)
- Fan Zhu (Presenter)
date: "`r Sys.Date()`"
output:
  pdf_document: default
fontsize: 11pt
geometry: margin=2cm
---

@@

For recoding date into a continuous variable, `date_num`, we filter out observations dated prior to the year 2010, as that is the year that the StreetRx web site was founded. For the seasonality variable, `date_quarter`, we extract the "quarter" of the year (numbered 1 through 4) as its own variable. Please see appendix for distributions.

@@

@@

We renamed this variable `dose_per_mg` to make its meaning clearer to the reader. This is a continuous variable that nevertheless tends to take on only a few values (please see appendix), as depicted in the histogram. We considered a recoding of this variable into levels, but decided against it as the dose per milligram technically may take on any value and the generalizability of our model may be improved if we are able to consider dose levels that are in a realistic range yet not present in our current data.

```{r}
# rename variable for ease of interpretability
df <- df %>% rename(dose_per_mg = mgstr)
```

@@

To consider whether we may want to include random slopes by state for our predictor variables, we examined plots of the predictor versus response among each state individually to determine whether it seemed that slopes fit to the group-level data would differ. (We present a sample of visualizations rather than visualizations for every state, due to space constraints. Please see appendix)

For our continuous variables (continuous date and dosage strength), in by-state plots we often saw strikingly similar patterns in predictor vs. response, suggesting that random slopes may not be necessary for these variables. For categorical variables, we generally observed was there was indeed variation among the states, but it was not clear whether this variation had any well-defined structure. It was not clear whether our models would benefit from random slopes for these variables, so we decided to further investigate this in our modeling process.

@@

# Appendix

## Code

```{r, ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}

## Visualizations

### Date variable  
```{r}
hist(df$date_num)
knitr::kable(table(df$date_quarter))
```

### Dose strength variable 
```{r}
hist(df$dose_per_mg)
```

### Region box plots
```{r}
ggplot(df, aes(x=reorder(USA_region, USA_region, length), y=log_ppm, fill=USA_region)) +
geom_boxplot() +
labs(title="log PPM by region", x="region",y="Log PPM")
```

### City visualization for FL
```{r}
ggplot(df[df$state=="Florida",], aes(x=log_ppm, y=reorder(city, city, length))) +
geom_boxplot() +
labs(title="log PPM by city (FL)", x="Log PPM",y="city")
```

### More EDA: slope and intercept distributions for predictors by state
```{r}
# To get a better sense of how state-level groups and slopes are distributed, we created the following intercepts for our coninuous variables:

@@

hist(group_int_list)
hist(group_slope_list)
```
